```{r}
# Set up
library(dplyr)
library(MASS)
library(ggplot2)

# Year 2021 to 2022
workforce = read.csv('UKWorkforce.csv', head = T)
ageyr21 = read.csv('Age_JulDec21.csv', head = T)
ageyr22 = read.csv('Age_JanJun22.csv', head = T)
age = rbind(ageyr21, ageyr22)
```

Hypothesis: How does doctor characteristics affect consumers' demand for dental healthcare across different age groups? 

Doctor characteristics: 
1. Gender (F/M)
2. Age (Under 35, 35 to 44, 45 to 54, 55+)
3. Contract Type (GDS, PDS, TDS, Mixed) ref: https://faq.nhsbsa.nhs.uk/knowledgebase/article/KA-01910/en-us
- GDS: It's necessary to meet the reasonable needs of its patients. This would exclude services provided for cosmetic, rather than health reasons - circumstances under which a treatment is necessary.
- PDS: Contracts are for specialist services such as sedation, domiciliary and generally have a time limit (they normally last for 5 years and are reviewed afterwards).
- TDS: Whereby the Commissioner (Sustainability and Transformation Partnership / Local Health Board) make payments, not the NHS Dental Services and these are usually emergency dental practices.

```{r}
# Data Preparation:
# workforce
workforce = subset(workforce, Year == '2021-22')
workforce = workforce[,-4]
workforce = subset(workforce, Parent_Code_1 != 'ENG')
workforce = workforce[,-1:-2]
workforce = aggregate(cbind(Dentist_Count) ~ Parent_Code_1 + Contract_Type + Age_Group + Gender + Dentist_Type, data = workforce, FUN = sum)
workforce = workforce[order(workforce$Parent_Code_1),]
write.csv(workforce, file = 'workforce.csv') # Export as workforce.csv to clean in excel
workforceT = read.csv('workforceT.csv', head = T)
workforce = workforceT # Workforce transposed dataset
workforce_density = read.csv('workforce_density.csv', head = T) # Workforce Density
```

workforceT.csv consists of number of dentists according to 
1. Gender (F/M)
2. Contract(GDS, PDS, TDS, MIXED)
3. Age (Under_35, X35to44, X45to54, X55)
4. Total number of dentists
5. Dentist type (Associate or Provider)

workforce_density.csv is the same as above but in terms of density

Rationale for regressing using workforce density: 
- To account for differences in dental sizes due to population differences in the regions 

```{r}
age = age[!is.na(age$POPULATION),]
age = subset(age, GEOG_TYPE == 'SUB_ICB')
age = age[,-2:-10]

# Splitting of age into age bands
age00 = subset(age, AGE_BAND == '0')
age00 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age00, FUN = sum)
age00$PATIENT_SEEN_DENSITY = age00$PATIENTS_SEEN / age00$POPULATION

age01 = subset(age, AGE_BAND == '1')
age01 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age01, FUN = sum)
age01$PATIENT_SEEN_DENSITY = age01$PATIENTS_SEEN / age01$POPULATION

age02 = subset(age, AGE_BAND == '2')
age02 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age02, FUN = sum)
age02$PATIENT_SEEN_DENSITY = age02$PATIENTS_SEEN / age02$POPULATION

age03 = subset(age, AGE_BAND == '3')
age03 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age03, FUN = sum)
age03$PATIENT_SEEN_DENSITY = age03$PATIENTS_SEEN / age03$POPULATION

age04 = subset(age, AGE_BAND == '4')
age04 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age04, FUN = sum)
age04$PATIENT_SEEN_DENSITY = age04$PATIENTS_SEEN / age04$POPULATION

age05 = subset(age, AGE_BAND == '5')
age05 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age05, FUN = sum)
age05$PATIENT_SEEN_DENSITY = age05$PATIENTS_SEEN / age05$POPULATION

age06 = subset(age, AGE_BAND == '6')
age06 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age06, FUN = sum)
age06$PATIENT_SEEN_DENSITY = age06$PATIENTS_SEEN / age06$POPULATION

age07 = subset(age, AGE_BAND == '7')
age07 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age07, FUN = sum)
age07$PATIENT_SEEN_DENSITY = age07$PATIENTS_SEEN / age07$POPULATION

age08 = subset(age, AGE_BAND == '8')
age08 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age08, FUN = sum)
age08$PATIENT_SEEN_DENSITY = age08$PATIENTS_SEEN / age08$POPULATION

age09 = subset(age, AGE_BAND == '9')
age09 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age09, FUN = sum)
age09$PATIENT_SEEN_DENSITY = age09$PATIENTS_SEEN / age09$POPULATION

age10 = subset(age, AGE_BAND == '10')
age10 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age10, FUN = sum)
age10$PATIENT_SEEN_DENSITY = age10$PATIENTS_SEEN / age10$POPULATION

age11 = subset(age, AGE_BAND == '11')
age11 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age11, FUN = sum)
age11$PATIENT_SEEN_DENSITY = age11$PATIENTS_SEEN / age11$POPULATION

age12 = subset(age, AGE_BAND == '12')
age12 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age12, FUN = sum)
age12$PATIENT_SEEN_DENSITY = age12$PATIENTS_SEEN / age12$POPULATION

age13 = subset(age, AGE_BAND == '13')
age13 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age13, FUN = sum)
age13$PATIENT_SEEN_DENSITY = age13$PATIENTS_SEEN / age13$POPULATION

age14 = subset(age, AGE_BAND == '14')
age14 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age14, FUN = sum)
age14$PATIENT_SEEN_DENSITY = age14$PATIENTS_SEEN / age14$POPULATION

age15 = subset(age, AGE_BAND == '15')
age15 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age15, FUN = sum)
age15$PATIENT_SEEN_DENSITY = age15$PATIENTS_SEEN / age15$POPULATION

age16 = subset(age, AGE_BAND == '16')
age16 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age16, FUN = sum)
age16$PATIENT_SEEN_DENSITY = age16$PATIENTS_SEEN / age16$POPULATION

age17 = subset(age, AGE_BAND == '17')
age17 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age17, FUN = sum)
age17$PATIENT_SEEN_DENSITY = age17$PATIENTS_SEEN / age17$POPULATION

age18to64 = subset(age, AGE_BAND == '18-64')
age18to64 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age18to64, FUN = sum)
age18to64$PATIENT_SEEN_DENSITY = age18to64$PATIENTS_SEEN / age18to64$POPULATION

age65to74 = subset(age, AGE_BAND == '65-74')
age65to74 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age65to74, FUN = sum)
age65to74$PATIENT_SEEN_DENSITY = age65to74$PATIENTS_SEEN / age65to74$POPULATION

age75to84 = subset(age, AGE_BAND == '75-84')
age75to84 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age75to84, FUN = sum)
age75to84$PATIENT_SEEN_DENSITY = age75to84$PATIENTS_SEEN / age75to84$POPULATION

age85 = subset(age, AGE_BAND == '85+')
age85 = aggregate(cbind(PATIENTS_SEEN, POPULATION) ~ PSEEN_END_DATE + REGION_CODE + REGION_ONS_CODE + REGION_NAME + PATIENT_TYPE + AGE_BAND, data = age85, FUN = sum)
age85$PATIENT_SEEN_DENSITY = age85$PATIENTS_SEEN / age85$POPULATION

```

```{r}
# Merging of datasets for modelling
# Dataset merged with workforce density
data00 = merge(age00, workforce_density, by=c('REGION_CODE')) # age00
data01 = merge(age01, workforce_density, by=c('REGION_CODE')) # age01
data02 = merge(age02, workforce_density, by=c('REGION_CODE')) # age02
data03 = merge(age03, workforce_density, by=c('REGION_CODE')) # age03
data04 = merge(age04, workforce_density, by=c('REGION_CODE')) # age04
data05 = merge(age05, workforce_density, by=c('REGION_CODE')) # age05
data06 = merge(age06, workforce_density, by=c('REGION_CODE')) # age06
data07 = merge(age07, workforce_density, by=c('REGION_CODE')) # age07
data08 = merge(age08, workforce_density, by=c('REGION_CODE')) # age08
data09 = merge(age09, workforce_density, by=c('REGION_CODE')) # age09
data10 = merge(age10, workforce_density, by=c('REGION_CODE')) # age10
data11 = merge(age11, workforce_density, by=c('REGION_CODE')) # age11
data12 = merge(age12, workforce_density, by=c('REGION_CODE')) # age12               
data13 = merge(age13, workforce_density, by=c('REGION_CODE')) # age13
data14 = merge(age14, workforce_density, by=c('REGION_CODE')) # age14
data15 = merge(age15, workforce_density, by=c('REGION_CODE')) # age15
data16 = merge(age16, workforce_density, by=c('REGION_CODE')) # age16
data17 = merge(age17, workforce_density, by=c('REGION_CODE')) # age17
data18 = merge(age18to64, workforce_density, by=c('REGION_CODE')) # age18to64
data19 = merge(age65to74, workforce_density, by=c('REGION_CODE')) # age65to74
data20 = merge(age75to84, workforce_density, by=c('REGION_CODE')) # age75to84
data21 = merge(age85, workforce_density, by=c('REGION_CODE')) # age85

# List of all datasets in one
datasets <- list(data00, data01, data02, data03, data04, data05, data06, data07, data08, data09, data10, data11, data12, data13, data14, data15, data16, data17, data18, data19, data20, data21)
```

Datasets to work with:
1. data00 to data21
2. datasets 

```{r}
# Data exploration
# Merging of datas into one
all_data <- rbind(data00, data01, data02, data03, data04, data05, data06, data07, data08, data09, data10, data11, data12, data13, data14, data15, data16, data17, data18, data19, data20, data21)
```

```{r}
# Data exploration
# Correlation matrix 
# Define the column indices to compute the correlation for
col_indices <- 9:21

# Loop over the datasets
for (i in 1:length(datasets)) {
  
  # Get the subset of the dataset by index
  dataset_subset <- datasets[[i]][, col_indices]
  
  # Compute the correlation matrix
  cor_mat <- cor(dataset_subset)
  
  # Correlation coefficients plot
  p <- ggplot() +
    geom_text(data = reshape2::melt(cor_mat), aes(x = Var1, y = Var2, label = round(value, 2), color = abs(value) > 0.7 & !is.na(value)), size = 3) +
    ggtitle(paste("Correlation matrix for dataset", i)) +
    scale_color_manual(values = c("black", "red")) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          axis.text.y = element_text(angle = 0, hjust = 1),
          plot.title = element_text(hjust = 0.5),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
  
  print(p)
  
  # Save plots - remove # in next line to save plots
  # ggsave(paste0("cor_matrix_", i, ".png"), plot = p, width = 8, height = 6, dpi = 300)

}

```

The results of the correlation matrix suggests that 
1. F_DENTIST and M_DENTIST are in perfect negative correlation
2. GDS and MIXED have a correlation coefficient of -0.7
3. PDS and TDS have a correlation coefficient of -0.7
4. TDS and Dentists aged under 35 have a correlation coefficient of 0.74. Although intuitively it does not make sense, removal of TDS due to reason 3 solves this issue.
5. Dentists aged 35 to 44 and Dentists aged 55+ have a correlation coefficient of -0.96
6. Dentists aged 35 to 44 and Dentists aged 45 to 54 have a correlation coefficient of 0.71
7. Dentists aged 45 to 54 and Dentists aged 55+ have a correlation coefficient of -0.84
8. Associate dentists and performing provider dentists are in perfect negative correlation.

Hence, we remove M_DENTIST, TDS, MIXED, Dentists aged 35 to 44, Dentists aged 55+ and ASSOCIATE from our model.

```{r}
# Data exploration
# Histogram of variables

# Main Dependent Variable - PATIENT_SEEN_DENSITY
ggplot(all_data, aes(x = PATIENT_SEEN_DENSITY)) +
  geom_histogram(color = "black", fill = "white") +
  labs(x = "Patient Seen Density", y = "Frequency") +
  theme_classic()

ggsave("PATIENT_SEEN_DENSITY_Histogram.png")

```

1. PATIENT_SEEN_DENSITY seems to be skewed with a larger left-tail. We might want to log the variable.
2. Since the density of dentists of different gender, contract types, age and dentists' types are homogenous throughout the time period in each regions, it makes no sense to plot the explanatory variables in histograms.

MODEL 1: log(PATIENT_SEEN_DENSITY) = F_DENTIST

```{r}
# Model 1: log(PATIENT_SEEN_DENSITY) = F_DENTIST

# Function that performs the linear regression and returns the results
perform_regression <- function(data) {
  # Linear Regression
  model <- lm(log(PATIENT_SEEN_DENSITY) ~ F_DENTIST, data=data)
  # Return the summary of the model
  return(summary(model))
}

summary_list <- list()

# Loop through the list of datasets and perform linear regression on each one
for (i in 1:length(datasets)) {
  # Perform regression on the current dataset and obtain the summary statistics
  summary <- perform_regression(datasets[[i]])
  summary_list[[paste0("Data", i)]] <- summary
}

# Save the summary list to a file for clarity
sink("summarystats_model1.txt")
cat("Summary Statistics for Linear Regression Models:\n\n")
cat("-------------------------------------------------\n\n")
for (i in 1:length(summary_list)) {
  # Print the dataset label
  cat(paste0("Data", i, ":\n"))
  # Print the summary of the model
  print(summary_list[[paste0("Data", i)]])
  # Add a blank line for separation
  cat("\n")
}

sink()
```

MODEL 2: log(PATIENT_SEEN_DENSITY) = GDS + PDS

```{r}
# Model 2: log(PATIENT_SEEN_DENSITY) = GDS + PDS

# Function that performs the linear regression and returns the results
perform_regression <- function(data) {
  # Linear Regression
  model <- lm(log(PATIENT_SEEN_DENSITY) ~ GDS + PDS, data=data)
  # Return the summary of the model
  return(summary(model))
}

summary_list <- list()

# Loop through the list of datasets and perform linear regression on each one
for (i in 1:length(datasets)) {
  # Perform regression on the current dataset and obtain the summary statistics
  summary <- perform_regression(datasets[[i]])
  summary_list[[paste0("Data", i)]] <- summary
}

# Save the summary list to a file for clarity
sink("summarystats_model2.txt")
cat("Summary Statistics for Linear Regression Models:\n\n")
cat("-------------------------------------------------\n\n")
for (i in 1:length(summary_list)) {
  # Print the dataset label
  cat(paste0("Data", i, ":\n"))
  # Print the summary of the model
  print(summary_list[[paste0("Data", i)]])
  # Add a blank line for separation
  cat("\n")
}

sink()
```

Results for MODEL 2 shows that GDS and PDS are generally statistically insignificant and may be considered to be removed from the final model.


MODEL 3: log(PATIENT_SEEN_DENSITY) = Under_35 + X45to54

```{r}
# Model 3: log(PATIENT_SEEN_DENSITY) = Under_35 + X45to54

# Function that performs the linear regression and returns the results
perform_regression <- function(data) {
  # Linear Regression
  model <- lm(log(PATIENT_SEEN_DENSITY) ~ Under_35 + X45to54, data=data)
  # Return the summary of the model
  return(summary(model))
}

summary_list <- list()

# Loop through the list of datasets and perform linear regression on each one
for (i in 1:length(datasets)) {
  # Perform regression on the current dataset and obtain the summary statistics
  summary <- perform_regression(datasets[[i]])
  summary_list[[paste0("Data", i)]] <- summary
}

# Save the summary list to a file for clarity
sink("summarystats_model3.txt")
cat("Summary Statistics for Linear Regression Models:\n\n")
cat("-------------------------------------------------\n\n")
for (i in 1:length(summary_list)) {
  # Print the dataset label
  cat(paste0("Data", i, ":\n"))
  # Print the summary of the model
  print(summary_list[[paste0("Data", i)]])
  # Add a blank line for separation
  cat("\n")
}

sink()
```

MODEL 4: log(PATIENT_SEEN_DENSITY) = PROVIDER 

```{r}
# Model 6: log(PATIENT_SEEN_DENSITY) = PROVIDER

# Function that performs the linear regression and returns the results
perform_regression <- function(data) {
  # Linear Regression
  model <- lm(log(PATIENT_SEEN_DENSITY) ~ PROVIDER, data=data)
  # Return the summary of the model
  return(summary(model))
}

summary_list <- list()

# Loop through the list of datasets and perform linear regression on each one
for (i in 1:length(datasets)) {
  # Perform regression on the current dataset and obtain the summary statistics
  summary <- perform_regression(datasets[[i]])
  summary_list[[paste0("Data", i)]] <- summary
}

# Save the summary list to a file for clarity
sink("summarystats_model4.txt")
cat("Summary Statistics for Linear Regression Models:\n\n")
cat("-------------------------------------------------\n\n")
for (i in 1:length(summary_list)) {
  # Print the dataset label
  cat(paste0("Data", i, ":\n"))
  # Print the summary of the model
  print(summary_list[[paste0("Data", i)]])
  # Add a blank line for separation
  cat("\n")
}

sink()
```

Account for interaction effects between age and gender. Dentist type should be irrelevant of age and gender.

FINAL MODEL: log(PATIENT_SEEN_DENSITY) = F_DENTIST + Under_35 + X45to54 + F_DENTIST*Under_35 + F_DENTIST*X45to54 + PROVIDER

```{r}
# FINAL MODEL: log(PATIENT_SEEN_DENSITY) = F_DENTIST+ Under_35 + X45to54 + F_DENTIST*Under_35 + F_DENTIST*X45to54

# Function that performs the linear regression and returns the results
perform_regression <- function(data) {
  # Linear Regression
  model <- lm(log(PATIENT_SEEN_DENSITY) ~ F_DENTIST + Under_35 + X45to54 + F_DENTIST*Under_35 + F_DENTIST*X45to54 + PROVIDER, data=data)
  # Return the summary of the model
  return(summary(model))
}

summary_list <- list()

# Loop through the list of datasets and perform linear regression on each one
for (i in 1:length(datasets)) {
  # Perform regression on the current dataset and obtain the summary statistics
  summary <- perform_regression(datasets[[i]])
  summary_list[[paste0("Data", i)]] <- summary
}

# Save the summary list to a file for clarity
sink("summarystats_finalmodel.txt")
cat("Summary Statistics for Linear Regression Models:\n\n")
cat("-------------------------------------------------\n\n")
for (i in 1:length(summary_list)) {
  # Print the dataset label
  cat(paste0("Data", i, ":\n"))
  # Print the summary of the model
  print(summary_list[[paste0("Data", i)]])
  # Add a blank line for separation
  cat("\n")
}

sink()
```

Age bands with significant results:
Age 0, 1, 2 and 18-64
Corresponding datasets: 
data00, data01, data02, data18
