Read data from csv-files

```{r}
#install.packages('MASS')
library(MASS)
BRC <- read.csv('data.csv', header=TRUE)
BRCprices<-as.matrix(BRC)

# Dates for which we have underlying and derivative data
Dates <-rev(as.Date(as.vector(BRCprices[,1]), "%m/%d/%y"))

# Underlying and derivative prices
BRCprices<-cbind(rev(as.numeric(as.vector(BRCprices[,2]))),
                 rev(as.numeric(as.vector(BRCprices[,3]))),
                 rev(as.numeric(as.vector(BRCprices[,4]))))
Derprices <- rev(as.numeric(as.vector(BRC[, 5])))

# Dates for which the option is active. We need stock path for all these
Trade_d <- read.csv('days.csv', header=TRUE)
Trade_days <- row.names(Trade_d)
Trade_days <-as.Date(Trade_days, "%m-%d-%y")

# Treasury bond yields
Yields <- cbind(rev(as.numeric(as.vector(BRC[,7]))),
                 rev(as.numeric(as.vector(BRC[,6]))),
                 rev(as.numeric(as.vector(BRC[,8]))))/100
```

Function for simulating multiple stock paths using gbm

```{r}
# GBM Function, simulate three paths, one starting at S0 + h, one starting at s0 and one starting at S0 - h depending on which underlying asset to vary
# This is to calculate greeks
library(MASS)
gbm.f<-function(s0,r,sigma,dt,m,underlying){ 
  p=length(s0)
  Sh=matrix(0,nrow=p,ncol=m+1)
  S_h=matrix(0,nrow=p,ncol=m+1)
  S=matrix(0,nrow=p,ncol=m+1)
  Sh[,1]=s0
  S_h[,1]=s0
  S[,1]=s0
  
  # Vary initial price of the underlying asset we wish to study greeks for
  h = 0.001*Sh[underlying,1] 
  Sh[underlying,1] = Sh[underlying,1] + h
  S_h[underlying,1] = S_h[underlying,1] - h
  
  # Important: simulate with same variables
  Z<-mvrnorm(m, r*dt-diag(sigma)*dt/2,sigma*dt)
  for(j in 2:(m+1)){
    Sh[,j]=exp(log(Sh[,j-1])+Z[j-1,])
    S_h[,j]=exp(log(S_h[,j-1])+Z[j-1,])
    S[,j]=exp(log(S[,j-1])+Z[j-1,])
  }
   # Exclude S0
  list("ph" = Sh[,2:(m+1)],"mh" = S_h[,2:(m+1)], "o" = S[,2:(m+1)])
}
```

Key stock data

```{r}
barrier1 = 2493.018
barrier2 = 6889.1295
barrier3 =  2347.5540
issue1 = 4532.76
issue2 = 12525.6900
issue3 = 4268.2800
coupon_dates <- as.Date(c("26-04-22", "26-07-22", "26-10-22", "26-01-23", "26-04-23", "26-07-23", "26-10-23", "22-01-24"), "%d-%m-%y") 
# Changed final to last trading day
redemption_dates <- as.Date(c("26-01-23", "26-04-23", "26-07-23", "26-10-23", "22-01-24"), "%d-%m-%y")

dt = 1/252
```

Payoff specification

```{r}
coupon_payment <- function(start_date, r) {
  coupon = 3.25/4
  # Discount factor
  v <- exp(r/4)
  discount_factors <- v ^ (7:0)
  cash_flow <- c(rep(coupon, 8))
  disc_cash_flows <- cash_flow * discount_factors
  if (start_date <= coupon_dates[1]) {
    cpo = sum(disc_cash_flows[1:8])
  } else if (start_date <= coupon_dates[2]) {
    cpo = sum(disc_cash_flows[2:8])
  } else if (start_date <= coupon_dates[3]) {
    cpo = sum(disc_cash_flows[3:8])
  } else if (start_date <= coupon_dates[4]) {
    cpo = sum(disc_cash_flows[4:8])
  } else if (start_date <= coupon_dates[5]) {
    cpo = sum(disc_cash_flows[5:8])
  } else if (start_date <= coupon_dates[6]) {
    cpo = sum(disc_cash_flows[6:8])
  } else if (start_date <= coupon_dates[7]) {
    cpo = sum(disc_cash_flows[7:8])
  } else if (start_date <= coupon_dates[8]) {
    cpo = disc_cash_flows[-1]
  }
  cpo
}

# Calculate pay-off at maturity of one path, given that we got in at a certain date
payoff <- function(S1, S2, S3, r, date) {
  min1 = min(S1)
  min2 = min(S2)
  min3 = min(S3)
  close1 = tail(S1, 1)
  close2 = tail(S2, 1)
  close3 = tail(S3, 1)
  if (close1 > issue1 & close2 > issue2 & close3 > issue3) {
    den = 100
  } else if (min1 < barrier1 | min2 < barrier2 | min3 < barrier3) {
    ret <- c(close1/issue1, close2/issue2, close3/issue3)
    den = min(ret)*100
  } else if (min1 == 0 | min2 == 0 | min3 == 0) {
    den = 0
  } else {
    den = 100
  }
  den + coupon_payment(date, r)
}
```

Redemption cost specification

```{r}
redemption <- function() {
  100 + 3.25/4
}
```


Functions for simulating path until maturity given the historical data (typically one year back in time)
Return full path during t = 0:T, which compromises historical data and simulated data
Also contains three different paths with varying starting points for the simulated part in order to calculate greeks

```{r}
sim_path <- function(S_historical, dates, date, r, underlying) {
  # dates = dates for historical data, date = current date
  n0=nrow(S_historical)
  Slogprices<-log(S_historical)
  Slogreturns<-Slogprices[2:n0,]-Slogprices[1:(n0-1),]
  sigma=cov(Slogreturns)/dt
  
  n = length(Dates)
  sindeces<-(1:n)["2022-01-18"<Dates & Dates<=date]
  # Number of historical dates
  nhist = length(sindeces)
  # Number of steps to simulate
  m=length(Trade_days) - nhist
  # Final historical date is starting point for simulation
  s0=S_historical[n0,]
  
  S<-gbm.f(s0,r,sigma,dt,m,underlying)
  # Extract path for different underlying assets and different starting points
  s1h=S$ph[1,]
  s2h=S$ph[2,]
  s3h=S$ph[3,]
  s1_h=S$mh[1,]
  s2_h=S$mh[2,]
  s3_h=S$mh[3,]
  s1=S$o[1,]
  s2=S$o[2,]
  s3=S$o[3,]

  # Put together historical data and future data
  Hists1<-BRCprices[sindeces,1]
  Hists2<-BRCprices[sindeces,2]
  Hists3<-BRCprices[sindeces,3]
  
  wholes1h<-c(Hists1,s1h)
  wholes2h<-c(Hists2,s2h)
  wholes3h<-c(Hists3,s3h)
  wholes1_h<-c(Hists1,s1_h)
  wholes2_h<-c(Hists2,s2_h)
  wholes3_h<-c(Hists3,s3_h)
  wholes1<-c(Hists1,s1)
  wholes2<-c(Hists2,s2)
  wholes3<-c(Hists3,s3)
  
  list("s1h" = wholes1h,"s2h" = wholes2h, "s3h" = wholes3h, "s1mh" = wholes1_h,"s2mh" = wholes2_h, "s3mh" = wholes3_h, "s1" = wholes1,"s2" = wholes2, "s3" = wholes3)
}

# Simulate n paths using the historical data, return paths from t = 0:T divided by underlying asset and starting point of simulated part
sim_n_path <- function(S_historical, dates, date, r, n_path, underlying) {
  # dates = dates for historical data, date = current date
  n = length(Dates)
  sindeces<-(1:n)["2022-01-18"<Dates & Dates<=date]
  nhist = length(sindeces)
  m=length(Trade_days) - nhist # Amount of steps to simulate is total minus historical
  S1h <- matrix(0, n_path, m + nhist)
  S2h <- matrix(0, n_path, m + nhist)
  S3h <- matrix(0, n_path, m + nhist)
  S1_h <- matrix(0, n_path, m + nhist)
  S2_h <- matrix(0, n_path, m + nhist)
  S3_h <- matrix(0, n_path, m + nhist)
  S1 <- matrix(0, n_path, m + nhist)
  S2 <- matrix(0, n_path, m + nhist)
  S3 <- matrix(0, n_path, m + nhist)
  for(i in 1:n_path){
    S = sim_path(S_historical, dates, date, r, underlying)
    # Extract path for different underlying assets and different starting points
    S1h[i,]= S$s1h
    S2h[i,]= S$s2h
    S3h[i,]= S$s3h
    S1_h[i,]= S$s1mh
    S2_h[i,]= S$s2mh
    S3_h[i,]= S$s3mh
    S1[i,]= S$s1
    S2[i,]= S$s2
    S3[i,]= S$s3
  }
  list("S1h" = S1h,"S2h" = S2h, "S3h" = S3h, "S1mh" = S1_h,"S2mh" = S2_h, "S3mh" = S3_h, "S1" = S1,"S2" = S2, "S3" = S3)
}
```

Function for simulating the price, delta and gamma at specified date using n_path samples. Note: BEFORE FIRST REDEMPTION DATE

```{r}
sim_n <- function(S_historical, dates, date, r, n_path, underlying, EMS, CV) {
  S = sim_n_path(S_historical, dates, date, r, n_path, underlying) # Simulate the paths
  
  # Extract based on starting points and asset
  S1h = S$S1h; S2h = S$S2h; S3h = S$S3h
  S1_h = S$S1mh; S2_h = S$S2mh; S3_h = S$S3mh
  S1 = S$S1; S2 = S$S2; S3 = S$S3
  
  # Perform Empirical Martingale correction of all paths
  if(EMS) {
    dateind = which(Trade_days == date)
    dateinds = dateind:length(Trade_days)
    S1h[,dateinds] = EMS.f(S1h[,dateinds], r[2]); S2h[,dateinds] = EMS.f(S2h[,dateinds], r[2]); S3h[,dateinds] = EMS.f(S3h[,dateinds], r[2]);
    S1[,dateinds] = EMS.f(S1[,dateinds], r[2]); S2[,dateinds] = EMS.f(S2[,dateinds], r[2]); S3[,dateinds] = EMS.f(S3[,dateinds], r[2]);
    S1_h[,dateinds] = EMS.f(S1_h[,dateinds], r[2]); S2_h[,dateinds] = EMS.f(S2_h[,dateinds], r[2]); S3_h[,dateinds] = EMS.f(S3_h[,dateinds], r[2]);
  }
  
  # Price option separately at current time for each starting point
  terminal_vh <- terminal_value(S1h, S2h, S3h, r[2], date) # Value at maturity
  terminal_v_h <- terminal_value(S1_h, S2_h, S3_h, r[2], date)
  terminal_v <- terminal_value(S1, S2, S3, r[2], date)
  Vh <- matrix(0, nrow = n_path, ncol = length(redemption_dates))
  V_h <- matrix(0, nrow = n_path, ncol = length(redemption_dates))
  V <- matrix(0, nrow = n_path, ncol = length(redemption_dates))
  Vh[,length(redemption_dates)] = terminal_vh
  V_h[,length(redemption_dates)] = terminal_v_h
  V[,length(redemption_dates)] = terminal_v
  
  # For simplicity, we assume we are always ahead of the first redemption date (january 2023)
  # At each redemption date, decide whether redemption happens or not for each path
  for(j in ((length(redemption_dates)-1):1)) {
    deltat = abs(as.numeric(difftime(redemption_dates[j+1],redemption_dates[j], units = "days"))/365) # Difference in years
    
    disc_vh = as.matrix(exp(-r[2]*deltat)*Vh[,j+1], ncol=1) # Discount previous values
    disc_v_h = as.matrix(exp(-r[2]*deltat)*V_h[,j+1], ncol=1)
    disc_v = as.matrix(exp(-r[2]*deltat)*V[,j+1], ncol=1)
    
    fh <- regress(S1h, S2h, S3h, redemption_dates[j], disc_vh, r[2]) # Model for calculating expected discounted payoff
    f_h <- regress(S1_h, S2_h, S3_h, redemption_dates[j], disc_v_h, r[2])
    f <- regress(S1, S2, S3, redemption_dates[j], disc_v, r[2])
    
    redcost <- redemption() # Cost of redemption
    for(k in (1:n_path)) {
      if(fh[k,1] <= redcost) {
        Vh[k,j] <- disc_vh[k,1] # If value less than redemption cost, issuer does not redeem
      } else {
        Vh[k,j] <- redcost # If value greater than redemption cost, issuer redeems
      }
      if(f_h[k,1] <= redcost) {
        V_h[k,j] <- disc_v_h[k,1] # If value less than redemption cost, issuer does not redeem
      } else {
        V_h[k,j] <- redcost # If value greater than redemption cost, issuer redeems
      }
      if(f[k,1] <= redcost) {
        V[k,j] <- disc_v[k,1] # If value less than redemption cost, issuer does not redeem
      } else {
        V[k,j] <- redcost # If value greater than redemption cost, issuer redeems
      }
    }
  }
  
  n0 = nrow(S_historical)
  h = 0.001*S_historical[n0,underlying]
  
  # Finally, discount to current day from first redemption date
  deltat = abs(as.numeric(difftime(redemption_dates[1], date, units = "days"))/365) # Difference in years
  Vcurrh <- exp(-r[2]*deltat)*Vh[,1]
  Vcurr_h <- exp(-r[2]*deltat)*V_h[,1]
  Vcurr <- exp(-r[2]*deltat)*V[,1]
  
  # Use control variates variance reduction to estimate discounted payoff or simple sample mean
  if(CV){
    # Use minimum of final/issue as control variate
    finalh <- matrix(0,n_path,3)
    finalh[,1] = S1h[,ncol(S1h)]/issue1;finalh[,2] = S2h[,ncol(S2h)]/issue2;finalh[,3] = S3h[,ncol(S3h)]/issue3
    CVsh <- as.vector(apply(finalh,1,min))
    payh = CV.f(Vcurrh,CVsh)$est
    
    final <- matrix(0,n_path,3)
    final[,1] = S1[,ncol(S1)]/issue1;final[,2] = S2[,ncol(S2)]/issue2;final[,3] = S3[,ncol(S3)]/issue3
    CVs <- as.vector(apply(final,1,min))
    p = CV.f(Vcurr,CVs)
    std <- p$std
    pay <- p$est
    
    final_h <- matrix(0,n_path,3)
    final_h[,1] = S1_h[,ncol(S1_h)]/issue1;final_h[,2] = S2_h[,ncol(S2_h)]/issue2;final_h[,3] = S3_h[,ncol(S3_h)]/issue3
    CVs_h <- as.vector(apply(final_h,1,min))
    pay_h = CV.f(Vcurr_h,CVs_h)$est
  } else {
    payh <- mean(Vcurrh); pay_h <- mean(Vcurr_h); pay <- mean(Vcurr) # Calculate payoffs by sample mean 
    std <- sd(Vcurr)
  }
  

  # Greeks by finite difference method
  delta = (payh - pay_h)/2*h
  gamma = (payh - 2*pay + pay_h)/h^2
  price = pay
  # non-callable price
  tau = abs(as.numeric(difftime(redemption_dates[5], date, units = "days"))/365) # Difference in years
  non_call_p = exp(-r[2]*tau)*mean(terminal_v)
  list("delta" = delta, "gamma" = gamma, "price" = price, "noncallp" = non_call_p, "std" = std)
}
```

Helper functions for simulating price at specified date

```{r}
# Calculates payoff at maturity for all paths
terminal_value <- function(S1, S2, S3, r, date) {
  terminal_v <- matrix(0,ncol = 1, nrow = nrow(S1))
  for(i in 1:nrow(S1)) {
    terminal_v[i,1] <- payoff(S1[i,], S2[i,], S3[i,], r, date)
  }
  terminal_v
}

# Model for calculating expected discounted payoff
regress <- function(S1, S2, S3, redemption_date, disc_v, r) {
  # Get data for regression
  prevdateinds <- 1:which(Trade_days == redemption_date)
  S1reg <- S1[,prevdateinds]
  S2reg <- S2[,prevdateinds]
  S3reg <- S3[,prevdateinds]
  # We use lowest fraction, as well as final prices as predictor variables
  low1 <- apply(S1reg,1,min)/issue1; low2 <- apply(S2reg,1,min)/issue2; low3 <- apply(S3reg,1,min)/issue3
  final1 <- S1reg[,ncol(S1reg)]; final2 <- S2reg[,ncol(S2reg)]; final3 <- S3reg[,ncol(S3reg)] 
  mlr <- lm(disc_v ~ final1 + final2 + final3 + low1 + low2 + low3 + final1^2 + final2^2 + final3^2 + low1^2 + low2^2 + low3^2 + final1^3 + final2^3 + final3^3 + low1^3 + low2^3 + low3^3)
  
  prediction = predict(mlr)
  tau <- abs(as.numeric(difftime(redemption_date, redemption_dates[5], units = "days"))/365)
  pos_payoff <- exp(-r*tau)*(100 + coupon_payment(redemption_date,r))
  
  # Set predictions higher than the maximum possible payoff to the maximum possible payoff
  for(i in (1:length(prediction))){
    if(prediction[i]>pos_payoff){
      prediction[i] = pos_payoff
    }
  }

  as.matrix(prediction, ncol=1) # Return predicted values 
}
```

Variance reduction functions

```{r}
EMS.f<-function(SimPaths,r){
  Ns=nrow(SimPaths); nsteps=ncol(SimPaths)-1
  Sstar=matrix(0,Ns,nsteps+1); W=matrix(0,Ns+1,nsteps)
  Sstar[,1]=SimPaths[,1]
  for(k in 2:(nsteps+1)){
    W[1:Ns,k-1]=Sstar[,k-1]*SimPaths[,k]/SimPaths[,k-1]
    W[Ns+1,k-1]=exp(-r*((k-1)*dt))*mean(W[1:Ns,k-1])
    Sstar[,k]=SimPaths[,1]*W[1:Ns,k-1]/W[Ns+1,k-1]
  }
  Sstar
}

# Control variates variance reduction. Returns estimate using the samples and corresponding control variates
CV.f<-function(Samples, CVs) {
  # First, find parameters needed for variance reduction using a tenth of all samples
  n_path = length(Samples)
  n_regress = ceiling(n_path/10)
  samples <- Samples[1:n_regress]
  variates <- CVs[1:n_regress]
  
  mu = mean(variates)
  c = - summary(lm(samples ~ variates))$coefficients[2]

  # Simulate price using 9/10 of all samples and variance
  sim_samples <- Samples[(n_regress+1):length(Samples)]
  sim_variates <- CVs[(n_regress+1):length(Samples)]
  est <- mean(sim_samples + c*(sim_variates - mu))
  std <- sd(sim_samples + c*(sim_variates - mu))
  list("est" = est, "std" = std)
}
```


Perform rolling pricing and sensitivity analysis of the option starting from 2 of may 2022 and 80 days in the future

```{r}
Nsim=10000

startingdate = "2022-05-02"
n_days = 80

n = length(Dates)-1
trainind<-(1:n)["2021-11-02"<Dates & Dates<= startingdate]
ntrain = length(trainind) 

initial_ind = which(Trade_days == startingdate)
price_days <- Trade_days[initial_ind:(initial_ind+n_days-1)]

prices <- numeric(n_days)
non_call_p <- numeric(n_days)
delta1 <- numeric(n_days)
delta2 <- numeric(n_days)
delta3 <- numeric(n_days)
gamma1 <- numeric(n_days)
gamma2 <- numeric(n_days)
gamma3 <- numeric(n_days)
sample_std <- numeric(n_days)

for(i in 1:n_days){
  date = price_days[i] # Get current date
  Yield = Yields[trainind[length(trainind)],]
  r = as.vector(Yield)
  print(r)

  # Delta and gamma for asset 1
  D1 = sim_n(BRCprices[trainind,], Dates[trainind], date, r, Nsim, 1, FALSE, FALSE) # First: EMS, second: CV
  delta1[i] = D1$delta
  gamma1[i] = D1$gamma
  # Delta and gamma for asset 2
  D2 = sim_n(BRCprices[trainind,], Dates[trainind], date, r, Nsim, 2, FALSE, FALSE)
  delta2[i] = D2$delta
  gamma2[i] = D2$gamma
  # Delta and gamma for asset 3
  D3 = sim_n(BRCprices[trainind,], Dates[trainind], date, r, Nsim, 3, FALSE, FALSE)
  delta3[i] = D3$delta
  gamma3[i] = D3$gamma
  
  # Use average to form price
  prices[i] = (D1$price + D2$price + D3$price)/3
  non_call_p[i] = (D1$noncallp + D2$noncallp + D3$noncallp)/3
  
  # Sample standard deviation of discounted payoffs
  sample_std[i] = sqrt(D1$std^2 + D2$std^2 + D3$std^2)/3
  
  # Update fitting data one step forward
  trainind=c(trainind[2:ntrain], trainind[ntrain]+1)
  print(i)
}
```

Visualize result of rolling simulation and calculate root mean square error

```{r}
dind1 = which(Dates == startingdate)
dinds = dind1:(dind1+n_days-1)

plot(prices~Dates[dinds], type="l" ,col="blue" , ylab="Price [%]" , xlab="", ylim=c(85,105))
abline(h=seq(80,105,5) , col="grey", lwd=0.8)
lines(Dates[dinds], Derprices[dinds], col="red")
lines(Dates[dinds], non_call_p, col="black")
legend(x = "topright",
       col = c("blue", "red", "black"), lty = 1, lwd = 1,
       legend = c('Simulated price', 'Actual Price', 'Simulated non-callable'))

mindelta <- min(min(delta1),min(delta2),min(delta3))
maxdelta <- max(max(delta1),max(delta2),max(delta3)) + 0.05
plot(delta1~Dates[dinds],type="l", col="blue", ylim=c(mindelta,maxdelta), ylab="Delta", xlab="")
abline(h=seq(mindelta,maxdelta, (maxdelta-mindelta)/5) , col="grey", lwd=0.8)
lines(Dates[dinds], delta2, col="red")
lines(Dates[dinds], delta3, col="black")
legend(x = "topright",
       col = c("blue", "red", "black"), lty = 1, lwd = 1,
       legend = c('Delta 1', 'Delta 2', 'Delta 3'))

mingamma <- min(min(gamma1),min(gamma2),min(gamma3))
maxgamma <- max(max(gamma1),max(gamma2),max(gamma3)) + 0.0005
plot(gamma1~Dates[dinds],type="l", col="blue", ylim=c(mingamma,maxgamma), ylab="Gamma", xlab="")
abline(h=seq(mingamma,maxgamma, (maxgamma-mingamma)/5) , col="grey", lwd=0.8)
lines(Dates[dinds], gamma2, col="red")
lines(Dates[dinds], gamma3, col="black")
legend(x = "topright",
       col = c("blue", "red", "black"), lty = 1, lwd = 1,
       legend = c('Gamma 1', 'Gamma 2', 'Gamma 3'))

RMSE = sqrt(sum((Derprices[dinds]-prices)^2)/length(prices))

cat("Root mean square error =", RMSE, "\n")
cat("Sample standard deviation =", mean(sample_std), "\n")
cat("Mean deltas:", mean(delta1), mean(delta2), mean(delta3), "\n")
cat("Mean gammas:", mean(gamma1), mean(gamma2), mean(gamma3), "\n")
```

Test volatility assumptions

```{r}

n = length(Dates)-1
d2021<-(1:n)["2021-01-01"<Dates & Dates<= "2021-12-31"]
d2022<-(1:n)["2021-12-31"<Dates]

S_historical1 <- BRCprices[d2021,]
n0=nrow(S_historical1)
Slogprices<-log(S_historical1)
Slogreturns<-Slogprices[2:n0,]-Slogprices[1:(n0-1),]
sigma1=cov(Slogreturns)/dt
print(sqrt(sigma1))

S_historical2 <- BRCprices[d2022,]
n0=nrow(S_historical2)
Slogprices<-log(S_historical2)
Slogreturns<-Slogprices[2:n0,]-Slogprices[1:(n0-1),]
Slogreturns<-Slogreturns[1:(n0-2),]
sigma2=cov(Slogreturns)/dt
print(sqrt(sigma2))
```

